---
title: "esm-244-lab-5-MW"
author: "Matthew Warham"
date: "2/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

library(tidyverse)
#install.packages("tseries")
#install.packages("forecast")
library(forecast)
library(tseries)

```

```{r data wrangle}

#US Residential energy consumption values (Jan 1973 - Oct 2017), Trillion BTU (US Energy Information Administration)

energy <- read_csv("energy.csv")
#View(energy)

```

c. Create 'ts' tiem series data ysing information in teh res_total column from energy

```{r}

res_ts <- ts(energy$res_total, frequency = 12, start = c(1973,1))

res_ts

plot(res_ts) # looks additive

```


d. Decompose ts

```{r decompose time series}

res_dc <- decompose(res_ts)
plot(res_dc)

#Plot Graphs
#1. Observed data
#2. moving average trend
#3. Seasonality
#4. Residuals

```

e. Other coolw ays to present time series data

```{r}

monthplot(res_ts)

#shows how each month has changed over months over entire series
# Increase over time during summer months

```


```{r}

ggseasonplot(res_ts)
ggseasonplot(res_ts, polar = TRUE) # puts all data for each year onto plot

```

f. Simple moving averages

```{r}

#forecast::ma()

sma_res <- ma(res_ts, order = 5) #changing window can hav a large impact on the data is viewed
plot(res_ts)
lines(sma_res, col = "red")

```

g. ACF (Autocorrelation)

```{r}

res_acf <- acf(res_ts)
#DRASTIC PEAKS every 12 months or every year

ggtsdisplay(res_ts)

```

i. Holt - Winters - Exponential smoothng methods

```{r}

res_hw <- HoltWinters(res_ts)
res_hw
plot(res_hw) #how models looks compared to actual data



```


Now some future forecasting with Holt Winters

```{r}

res_forecast <- forecast(res_hw, h = 60)
plot(res_forecast)

```

Check Distribution of the residuals

```{r}

hist(res_forecast$residuals)

```


Autoregressive integrated moving average (ARIMA) using auto.arima for p,d,q

```{r}

res_pdq <- auto.arima(res_ts)
res_pdq

#ARIMA(1,0,2)(0,1,1)[12] with drift



```


pdq: [1,0,2], [0,1,1]

Fit ARIMA model using thos epdq values

```{r}

res_arima <- arima(res_ts, order = c(1,0,2), seasonal = list(order = c(0,1,1)))
res_arima

# Check out he residuals
par(mfrow = c(1,2))
hist(res_arima$residuals)
qqnorm(res_arima$residuals)

```


Now forecast with ARIMA outputs

```{r}

forecast_res <- forecast(res_arima, h = 72)
plot(forecast_res)

```

Now view this in ggplot

```{r}

res_df <- data.frame(forecast_res)
month_seq <- seq(1,72)

res_df_2 <- data.frame(month_seq, res_df)

#view(res_df_2)

ggplot(res_df_2, aes(x = month_seq, y = Point.Forecast))+
  geom_line()

```

